cpcrw_ms

### 2016-01-31 manuscript/cpcrw_manuscript.Rmd
Filled out and tweaks methods text

The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (http://www.lter.uaf.edu/default.cfm). Annual average air temperature was 􏰀3.3°C, and the average temperature in July was 16.4°C, which was the maximum in this watershed. Annual average precipitation was 269 mm, and the average precipitation of 47 mm in August was the highest in this area.

SE slope, transition zone...

of this study site was more than TODO years according to the tree rings at stem base. Stand density of dominant Picea mariana and glauca trees (>1.3 m in height) was TODO trees ha^-1^, with a basal area of TODO m^2^ ha^-1^.

@@

In the lab, the soil cores were stored at 4 &deg;C for several days until they were weighed and prepared for incubation. At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (each N=6). These groups included two incubation temperatures (4 &deg;C and 20 &deg;C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume. The fifth group was a 20 &deg;C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 &deg;C "drought" group. The final group includedsix cores for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density. Subsamples were collected and stored at -20 &deg;C for potential enzyme activity and ATP.

On 18 August 2015 cores were placed into one of two growth chambers (Conviron CMP6050 Control Systems, Winnipeg, Canada) maintained at 4 and 20 &deg;C temperatures and 70% relative humidity and allowed to equilibrate for two weeks. Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) evolution four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation. Throughout the incubation, cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action and to moderate the drying of the 20 &deg;C "controlled drought" group. To maintain the field moisture treatment groups at ~28% moisture, the cores sat on saturated pore plates and received additional wetting from the top of the cores to supplement for when evaporation exceeded imbibition.

For each measurement, a six-core treatment group was connected to a Picarro A0311 multiplexer that was in turn connected to a Picarro G2301 GHG Analyzer (Picarro Inc., Santa Clara, CA, USA). Dry GHG concentration were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group. Ambient air was measured between treatment groups, and before starting measurements in a chamber.

@@

AOutlier exclusion based on 3.2 mean absolute deviations from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.

>>>>

The field component of this research took place in Caribou-Poker Creeks Research Watershed (CPCRW), part of the Bonanza Creek LTER (http://www.lter.uaf.edu/default.cfm). Annual average air temperature was -2.5 °C, and annual average precipitation was 400 mm {Petrone, 2006 #4313}. The watershed's lowlands and north-facing slopes are dominated by black spruce (*Picea mariana*), feathermoss (*Pleurozium schreberi* and others), and *Sphagnum* spp.; the drier south slopes tend to be deciduous with a mixture of quaking aspen (*Populus tremuloides*), paper birch (*Betula neoalaskana*), and patches of alder (*Alnus crispa*).

We sampled soils from a southeast slope (65.1620 °N, 147.4874 °W), in a 60 m transition zone between lowland *Picea mariana* Mill BSP. and upland *Betula neoalaskana*, with significant white spruce (*Picea glauca*) presence. Stand density in this transition zone was 4060 ± 2310 trees ha^-1^, with basal area of 27.9 ± 7.0 m^2^ ha^-1^. The forest was at least 85 years old according to tree rings at the stem base.

@@

In the lab, the soil cores were stored at 4 °C for several days until they were weighed and prepared for incubation. At that point (11-12 August 2015), three fragmented or otherwise damaged cores were discarded, and the remaining cores were randomly assigned to one of six groups (N=6 in each group). These groups included two incubation temperatures (4 °C and 20 °C) each with two moisture treatments: one in which soil moisture was maintained at field conditions (~28% moisture by volume), and a drought treatment in which no water was added and cores were allowed to dry down to ~5% moisture by volume. The fifth group was a 20 °C "controlled drought" one, in which water was added so that these cores' moisture status closely matched those of the 4 °C "drought" group. The final 6-core group was used for destructive, pre-incubation measurements including moisture content, pH, soil carbon and nitrogen, and bulk density. Subsamples were collected and stored at -20 °C for potential enzyme activity and ATP.

On 18 August 2015 cores were placed into one of two growth chambers (Conviron CMP6050 Control Systems, Winnipeg, Canada) maintained at 4 and 20 °C temperatures and 70% relative humidity and allowed to equilibrate for two weeks. Starting on 31 August 2015 we measured the cores' mass and GHG (CH~4~ and CO~2~) evolution four times in the first week, then twice per week for the first month, and then once per week for the rest of the 100-day incubation. Throughout the incubation, cores were mounted on 0.05-MPa ceramic plates (Soil Moisture Equipment Corp., Santa Barbara, CA, USA) so that, when the plates were placed in contact with water, water would move up into the cores via capillary action. This was done to moderate the drying of the 20 °C "controlled drought" group. The 4 and 20 °C "field moisture" groups received additional wetting from the top of the cores to maintain their water status.

For each measurement, a six-core treatment group was connected to a Picarro A0311 multiplexer that was in turn connected to a Picarro G2301 GHG Analyzer (Picarro Inc., Santa Clara, CA, USA). Dry CH~4~ and CO~2~ concentrations were monitored for 2 minutes, and this was repeated 2-3 times before moving on to a new treatment group. Cores were weighed immediately after (PEYTON) gas measurements. Ambient air was measured between treatment groups, and before starting measurements in a chamber. The incubation experiment concluded on 10 December 2015.

@@

Outlier exclusion based on 3.2 mean absolute deviations from treatment and temperature mean. We excluded X of Y (Z%) of the summarized samples for this reason.
